\textbf{Teaching a machine to generate an artificial universe}

### 2020-09-02 paper.tex
new test figs, a bit more text

\twocolumn[
	\begin{@twocolumnfalse}
		\maketitle
		\begin{abstract}
		    This work presents a three-dimensional generative framework that uses an Auxiliary Classifier Generative
		    Adversarial Network to create realistic samples of cosmological structure at various values of redshift~$z$. 
		    Provisional samples of the virtual cosmic web are obtained from N-body simulation data that is restricted to 
		    volumes and projections of size $200 \, h^{-1} \, \text{Mpc}$. The generated samples are shown to demonstrate
		    a high degree of coherence with their simulated counterparts and the distinction is difficult to make visually
		    by human experts. The nature of the separation between generated and N-body simulation samples is quantified 
		    with repeated tests of the Kullback-Leibler divergence on ensembles of samples alongside samples with 
		    uniformly distributed pixel densities. It is found that the implementations secure a coherence to within the
		    same order of magnitude as a self comparison of the simulation samples. The main advantage of the generative
		    approach to simulating large volumes of cosmological structure is that much less time is required than 
		    creating N-body simulations on high performance computing resources. This is demonstrated by the creation of
		    large box volumes made with $\sim10^3$ generated samples for a redshift $z$ arranged on a cubic mesh. At this 
		    stage, the generative artificial intelligence can then be said to be capable of generating an artificial 
		    universe. Future explorations of the method that are relevant to the next paradigm in cosmological simulations 
		    are also outlined. 
		  %  The most pertinent application being for a generator that can produce samples that maintain high statistical 
		  %coherence to simulation data at various redshifts.

            %\textit{}

		\end{abstract}
	\end{@twocolumnfalse}
]
{\renewcommand{\thefootnote}%
    {\fnsymbol{footnote}}
  \footnotetext[1]{Corresponding author: \url{jedhmr@gmail.com}}
}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/diagrams/GAN_diagram1.png}
\centering
\caption{Schematic of the GAN framework; including discriminator and generator networks, the latent-space vector, dataset 
of samples and the backpropagation feedback to model parameters.}
\label{fig:GAN_diagram}
\end{figure}

@@

The type of generative model chosen for the objective of this work was the Generative Adversarial Network (GAN) invented 
by Goodfellow et al. 2014~\cite{gf_gan}. The successes of GANs are well established in a wide range of applications on 
natural images~\cite{wgan, karrasgan, largegan} and there are many extensions to GANs that open up a wide range of other
successful models~\cite{pix2pix, hiresgan, lapgan}.

% BASICS: WHAT DO D AND G DO?
Generative Adversarial Networks consist of two models known as the generator and the discriminator. The generator must 
create samples that are allegedly drawn from the same distribution as the real samples dataset. The discriminator assigns 
scalars to a collection of input data and generated samples that represent the probability that each sample is real and 
from the data distribution $p_{data}$ or fake and from the model distribution  $p_{model}$. The discriminator has access 
to both real and fake samples. The training of a GAN involves navigating the combined parameter space of the generator 
and discriminator network parameters to find the optimal value of a function that gives the best performance of the 
networks at their opposing objectives. The metric that describes the performance is related to the distance between 
the model and data distributions. There is no need for a validation subset external to the training data because the 
generator does not have direct access to the training data. This is not the case for singular deep learning networks.

@@

% WHAT ARE D AND G? HOW DO D AND G WORK?
The discriminator and generator models can be represented as functions $D$ and $G$ respectively because of their 
mappings as neural networks and parameter dependencies. The discriminator $D$ has input $\mathbf{x}$ and parameters 
$\bm{\theta}^{(D)}$. The generator $G$ takes a latent space vector $\mathbf{z}$ as input and uses parameters 
$\bm{\theta}^{(G)}$. Information flows through the network and its parameters to create the outputs $D(\mathbf{x})$ or 
$D(G(\mathbf{z}))$ for the discriminator and $G(\mathbf{z})$ for the generator. This is the forward-propagation from 
an input to an output. The model performances are calculated with the objective functions $J^{(D)}$ or $J^{(G)}$.

@@

in the combined parameter space. This equation shows the opposition between the maximization of the correct 
classifications by $D$ and the minimisation of the classifications by $D$ of the generator products that are made 
by $G$.

@@

This motivates the generator to increase the likelihood of the discriminator incorrectly classifying a sample. The 
discriminator tries to increase its own likelihood of correctly classifying a sample by minimising the binary 
cross-entropy objective. The local minima of the value function $V$ are not necessarily Nash equilibrium points, rather
they are points that simultaneously minimise the objective functions $J^{(D)}$ and $J^{(G)}$. Whilst both networks search 
the parameter space via SGD of the value function, there is no guarantee of a Nash equilibrium, this is part of what makes 
training an adversarial network a difficult task.

@@

\subsection{Auxiliary-Classifier GANs}
The ability to generate different classes of an object does not require separate and independent GANs. The GAN framework 
may be supplemented by an auxiliary classifier model that is attached to the discriminator model, which still classifies 
the samples as real and fake, whilst additionally characterising the pre-designated class of the object. This augmented 
GAN was derived by Odena et al. 2017 \cite{acgan} and is referred to as an Auxiliary Classifier Generative Adversarial 
Network (ACGAN). The application to this method is that the different redshift data available from N-body simulations can 
be classified and given to an ACGAN. The generator component of this model can create samples at different redshifts. 
These samples could be tessellated together to form a larger distribution across a continuous range of  redshifts. A very 
similar method known as a Conditional-GAN (CGAN)~\cite{cgan} is known to work well at the same task.

In the ACGAN framework each generated sample has a corresponding label $c \sim p_c$ and noise vector $\mathbf{z} \sim p_z$ which 
make the generator create a sample $G(c, \mathbf{z})$. The discriminator $D$ gives probability distributions $P(S|X)$ and $P(C|X) 
= D(X)$ over the sources of the samples $S$ and class labels $C$. The objective function is made of two parts; the 
likelihood of a correct source $L_S$ and the log-likelihood of the correct class $L_C$

\begin{eqnarray}
    L_S &=& \EE [\log P(S = \textup{real} | X_{real})] \nonumber \\
        &+& \EE [\log P(S = \textup{fake} | X_{fake})]
\end{eqnarray}

\begin{eqnarray}
    L_C &=& \EE [\log P(C = c | X_{real})] \nonumber \\
        &+& \EE [\log P(C = c | X_{fake})]
\end{eqnarray}

In this case $D$ tries to maximize $L_S + L_C$ and $G$ tries to minimise $L_C + L_S$.

@@

@@

\subsection{Fourier Transform ACGAN (FT-ACGAN)}
% write this if I get it to work

@@

The galaxy catalogs containing physical properties of galaxies in the Millennium simulation were derived with the 
Semi-Analytic Galaxy Evolution (SAGE) codebase by Croton et al. 2016~\cite{Croton2016} that improves on the earlier 
model of Croton et al. 2006~\cite{Croton2006}. The SAGE catalogs were sourced from the Theoretical Astrophysical 
Observatory (TAO)~\cite{TAO}. SAGE modelling adds baryonic processes into N-body simulations of halos after the 
simulations have been run. The baryonic particle properties are from information in the simulation run such as mass, 
size, spin and their merger history. SAGE models account for different mechanisms that affect the evolutions of 
galaxies and therefore the evolution of cosmological structure. The N-body particle positions of the Millennium-SAGE 
data were taken at snapshots at discrete redshifts $z=\{0.0, 1.07, 2.07, 3.06, 4.18, 5.28\}$ to give the training 
data for the ACGAN.

@@

\subsection{Extraction of training data} 
% TRAINING DATA AND PREPROCESSING
Position vector components for the centres of cube samples in the simulation volume were drawn from a uniform distribution 
along each axis of each redshift-snapshot. If the addition or subtraction of half the side length of the cube to any of 
the components resulted in a point outside of the simulation volume, the point was discarded. The galaxy cube samples were 
obtained by sorting the position vectors of galaxies in the simulation that were within the cube volume. Each point inside 
the cube was recorded in an array. From each simulation volume of $512 \: h^{-1} \,\textup{Mpc}$ ensembles of 64 array 
samples of size $80 \: h^{-1} \,\textup{Mpc}$ were selected. The galaxy positions in each array for every cube were scaled
to the range $[-1,+1]$. 3D-histograms with 32 bins per axis were made on each cube across the same range for each axis. 
The densities in each bin cell were also scaled to this range to fit the output ranges of the GAN models. 
% After testing different sizes, a cube size of $160 \: h^{-1} \,\textup{Mpc}$ was found to balance the selection of unique 
% samples in the simulation volume against featuring the same features in the simulation structure.

\begin{figure}
\includegraphics[width=\columnwidth]{figures/diagrams/tt_scaled.png}% scaling2.png
\centering
\caption{The application of the scaling $s$ to a 3D-histogram training sample. The cube on the left shows an 
image $x$ and on the right is a scaled image $s(x)$.}
\label{fig:scaling}
\end{figure}

@@

\subsection{Initial data transformations} 
The 3D-histograms have pixel densities that span across many orders of magnitude from voids to dense clusters of galaxies 
for each redshift value. The origin of GANs from image generation tasks has optimised them for the general statistical 
structure of these images - the pixel densities in natural images are typically more uniformly distributed over the same 
scales. To help the GAN in generating sparsely distributed densities a transform, first utilised in~\cite{web_gan}, was 
used on the histograms to enhance the contrast between the density features. The transform from the original histogram 
$x$ to the scaled histogram $s(x)$ is

\begin{equation}
    s(x) = \frac{2x}{x+a} - 1
\end{equation}

where $a$ is a free parameter that depends on the sparsity of the histogram densities. The value of $a=1.25$ was found 
to be optimal for the $160 \:  \: h^{-1} \,\textup{Mpc}$ histograms at all redshifts $z$. This parameter remains fixed 
during training. The value of $a$ controls the median pixel density of the scaled image. Figure \ref{fig:scaling} shows 
the difference between scaled images $s(x)$ and the original images $x$. The inverse transformation $s^{-1}(x)$ can be 
applied to obtain the original image. 
% A logarithmic scaling was used originally, but this did not provide good results with the GAN architecture that used 
% the hyperbolic tangent ($\tanh$) activation because the logarithm of a negative number is undefined. The models using 
% $\tanh$ activations achieved much better results than models without these activations.

@@

\textit{Architecture} --- 
Table~\ref{table:3dgan_archi} of Appendix~\ref{appendix:gan_setups} shows the architectures of the ACGAN discriminator 
and generator models. The hyper-parameters used in the ACGAN are also given in Table~\ref{table:acgan_params} of 
Appendix~\ref{appendix:acgan_setups}. Together the discriminator and generator models had around $2.5\times10^8$ trainable 
parameters across one linear layer and three convolutional layers for the generator model and three convolutional layers 
for the discriminator model. \\

@@

The ACGAN implementation was limited by the GPU memory capacity - the parameters and architectures together with the 
sample batch sizes meant that the GPU was easily saturated. Deeper models with more convolutional units were implemented 
but these tended to stall the GPU. A training loop for $10^5$ epochs for the $200 \: h^{-1}\: \textup{Mpc}$ cubes 
formatted to $32 \times 32 \times 32$ pixel histograms took approximately 72 hours on an NVIDIA Tesla V100 GPU with 
32GB capacity.

@@

% {\setlength{\parindent}{0cm}
% \textit{Redshift interpolation} --- 
% To observe the A fixed point in the latent space is chosen and 
% }
\subsubsection{Redshift interpolation}
% talk about what this is in our method, what comes out of doing it 
The latent space of a generator in a typical GAN framework stores a compressed representation of the statistical structure
of the distribution of the real data. By choosing different points in this space and supplying them to the generator, 
samples can be derived with different characteristics that have been extracted from the training process of the discriminator
and generator models.

@@

An array of latent-space vectors and class labels of a given redshift are passed to the trained ACGAN generator to create a
collection of cube samples. These are placed together face-to-face to create larger samples of cosmic structure.

@@

\subsubsection{Latent space navigation}
% show the maths behind it, the general method of doing it. See https://arxiv.org/pdf/1609.04468.pdf, % (Larsen et al., 2016).
In all non-trivial cases, the number of parameters available to the generative model of a GAN through its architecture is less than the total parameters that constitute the training set. This means the discriminator and generator must find 
condensed representations of the training data. The generator model is sampled from a set of latent variables in the latent
space. Vector space arithmetic with points in the latent space allows interpolation between tenable examples of learned 
representations of the training data the GAN is exposed to in training. 
% With the inherent structuring of the latent space by GANs, these operations can allow complex

Interpolation is the transit between two distinct points in the latent space. In this work, spherical interpolation is used
as opposed to linear interpolation. The latter being a over-simplistic approach as the latent space of the model trained in this work had 256 dimensions with a Gaussian prior $\mathcal{N}(0,1)$. Spherical interpolation uses a great circle arc on the N-dimensional hypersphere in the latent space, with a path given by % (Shoemake 85)

where a discrete set of points between points $q_1$ and $q_2$ are passed to the generator model to create samples with.

\subsubsection{Building larger volumes from generated samples}
% say how the larger volumes are made from smaller generated cubes, talk about 
A selection of cubes were generated with a batch of $N$ latent vectors $\mathbf{Z}=\{\mathbf{z}_i, ..., \mathbf{z}_N\}$ drawn from the same prior $\mathcal{N}(0,1)$ as the generative model in training. These are placed at points on a cubic mesh to create a larger box of generated cubes. This was repeated for side-lengths 7, 9 and 11 (in multiples of the initial generated cube side-lengths).

@@

\subsection{Comparison of distributions with Kullback-Leibler Divergences}
Generative Adversarial Networks reduce the difference of the model probability distribution between the generated samples 
$p_{model}$ and the probability distribution of the training data samples $p_{data}$. This reduction quantifies how well 
$p_{model}$ estimates $p_{data}$. An estimate on the suitability of this replacement was made using a comparison with 
different ensembles of samples. The ensembles were collected from the training, generated and uniform distributions of 
histograms denoted $t$, $g$ and $u$. The uniform distribution $u$ was a set of histograms with pixel densities drawn from 
a uniform distribution $\mathcal{U}(-1,1)$. The comparison of sample ensembles tested how well the generator made the 
discriminator associate the real and generated samples to the same distribution. The Kullback-Leibler (KL) divergence and 
its relevance to the comparisons is defined in Appendix~\ref{appendix:kl_div}.

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/kernel_fig.png}
\centering
\caption{A histogram down-sampled with three kernels of size $k=\{1, 2, 4\}$ (from left to right).}
\label{fig:kernel_fig}
\end{figure}

@@

\subsubsection{Deriving the KL comparative diagnostic}
There were 64 samples in the ensembles from the distributions $t$, $g$ and $u$ that contained 128 samples each. Each 
histogram in the ensemble was down-sampled by iterating a kernel of size $k$ over each histogram and summing the densities. 
The summed densities were divided by the number of bins in the kernel after down-sampling. This is equivalent to the 
initial histograms having a lower sampling with larger bins. After an ensemble is down-sampled, the ensembles of two 
distributions are compared. The samples are individually flattened to `histogram-vectors' with $(32/k)^3$ components. 
This is shown in Figure~\ref{fig:kernel_fig}. One-dimensional histograms with two bins were made for each component in the 
histogram-vectors. The two bins represented `high' and `low' pixel densities and the boundary between them was the median 
of the training data pixel densities. This is because using the median meant that it was not assumed that the pixel density 
distributions were symmetric. The histogram-vectors were then collapsed over the ensemble and normalised to sum to one as 
a probability distribution. The vector was then sampling the probability density function of pixel density. This process 
is illustrated in Figure \ref{fig:klstat}.

@@

\subsection{Generated images}
% show generated images at various redshifts, compare with training data

\begin{figure*}[hbt!]
\includegraphics[width=15cm]{figures/cubes/classes3d.jpg}
\centering
\caption{A collection of training data cubes extracted from the Millennium-SAGE catalogs. Each row corresponds to the redshift value given above each cube. The scaling $s(x)$ is applied to each cube $x$.}
\label{fig:gan_game}
\end{figure*}

@@

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/mill_box_3.png}
\centering
\caption{A box made of generated cubes at redshift $z=3.19$}
\label{fig:kernel_fig}
\end{figure}

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/gen_box_9_z3.png}
\centering
\caption{A box made of generated cubes at redshift $z=3.0$.}
\label{fig:kernel_fig}
\end{figure}

@@

\subsection{Loss and $D_{KL}$ curves}

>>>>

\twocolumn[
	\begin{@twocolumnfalse}
		\maketitle
		\begin{abstract}
		    This work presents a three-dimensional generative framework that uses an Auxiliary Classifier Generative
		    Adversarial Network to create realistic samples of cosmological structure at various values of redshift~$z$. 
		    Provisional samples of the virtual cosmic web are obtained from N-body simulation data that is restricted to 
		    volumes of size $200 \, h^{-1} \, \text{Mpc}$. The generated samples are shown to demonstrate
		    a high degree of coherence with their simulated counterparts and the distinction is difficult to make visually
		    by human experts. The nature of the separation between generated and N-body simulation samples is quantified 
		    with repeated tests of the Kullback-Leibler divergence on ensembles of samples alongside samples with 
		    uniformly distributed pixel densities. It is found that the implementations secure a coherence to within the
		    same order of magnitude as a self comparison of the simulation samples. The main advantage of the generative
		    approach to simulating large volumes of cosmological structure is that much less time is required than 
		    creating N-body simulations on high performance computing resources. This is demonstrated by the creation of
		    large box volumes made with $\sim10^3$ generated samples for a redshift $z$ arranged on a cubic mesh. At this 
		    stage, the generative artificial intelligence can then be said to be capable of generating an artificial 
		    universe. Future explorations of the method that are relevant to the next paradigm in cosmological simulations 
		    are also outlined. 
		  %  The most pertinent application being for a generator that can produce samples that maintain high statistical 
		  %coherence to simulation data at various redshifts. 
            %\textit{}
		\end{abstract}
	\end{@twocolumnfalse}
]
{\renewcommand{\thefootnote}%
    {\fnsymbol{footnote}}
  \footnotetext[1]{Corresponding author: \url{jedhmr@gmail.com}}
}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/diagrams/ACGAN_diagram_box.png}
\centering
\caption{Schematic of the GAN framework; including discriminator and generator networks, the latent-space vector, dataset 
         of samples and the backpropagation feedback to model parameters. \textit{Source} and \textit{Class} refer to the
         values of $p_\mathbf{x}$ and $p_\mathbf{c}$ designated by the discriminator on a sample $\mathbf{x}$.}
\label{fig:GAN_diagram}
\end{figure}

@@

The type of generative model chosen for the objective of this work was the Auxiliary Classifier Generative Adversarial 
Network (ACGAN) created by Odena et al. 2017~\cite{acgan} which is an extension to the original GAN invented by Goodfellow 
et al. 2014~\cite{gf_gan}. The successes of GANs are well established in a wide range of applications on natural 
images~\cite{wgan, karrasgan, largegan} and there are many extensions to GANs that open up a wide range of other successful models~\cite{pix2pix, hiresgan, lapgan}.

% BASICS: WHAT DO D AND G DO?
Generative Adversarial Networks consist of two models known as the generator and the discriminator. The generator must 
create samples that are allegedly drawn from the same distribution as the real samples dataset. The discriminator assigns 
scalars to a collection of input data and generated samples, each representing the probability that a sample is either real 
and from the data distribution $p_{data}$ or fake and from the model distribution  $p_{model}$. The discriminator has 
access to both real and fake samples. The training of a GAN involves navigating the combined parameter space of the 
generator and discriminator network parameters to find the optimal value of a function that gives the best performance of
the networks at their opposing objectives. The metric that describes the performance is related to the distance between 
the model and data distributions. There is no need for a validation subset external to the training data because the 
generator does not have direct access to the training data. This is not the case for singular deep learning networks.

@@

% WHAT ARE D AND G? HOW DO D AND G WORK?
The discriminator and generator models can be represented as functions $D$ and $G$ respectively because of their 
mappings as neural networks and parameter dependencies. The discriminator $D$ has image input $\mathbf{x}$ and parameters 
$\bm{\theta}^{(D)}$. The generator $G$ takes a latent space vector $\mathbf{z}$ as input and uses parameters 
$\bm{\theta}^{(G)}$. Information flows through the network and its parameters to create the outputs $D(\mathbf{x})$ or 
$D(G(\mathbf{z}))$ for the discriminator and $G(\mathbf{z})$ for the generator. This is the forward-propagation from 
an input to an output. The model performances are calculated with the objective functions $J^{(D)}$ or $J^{(G)}$.

@@

in the combined parameter space. This equation shows the opposition between the maximization of the correct 
classifications by $D$ and the minimisation of the correct classifications by $D$ of the generator products that are 
made by $G$.

@@

This motivates the generator to increase the likelihood of the discriminator incorrectly classifying a sample. The 
discriminator tries to increase its own likelihood of correctly classifying a sample by minimising the binary 
cross-entropy objective. The local minima of the value function $V$ are not necessarily Nash equilibrium points, rather
they are points that simultaneously minimise the objective functions $J^{(D)}$ and $J^{(G)}$. Whilst both networks search 
the parameter space via SGD of the value function, there is no guarantee of a Nash equilibrium, this is part of what makes 
training an adversarial model such a difficult task.

@@

\subsection{Auxiliary-Classifier GANs}
The ability to generate different classes of an object does not require separate and independent GANs. The GAN framework 
may be supplemented by an auxiliary classifier model that is attached to the discriminator model, which still classifies 
the samples as real and fake, whilst additionally characterising the pre-designated class of the object. This augmented 
GAN was derived by Odena et al. 2017 \cite{acgan} and is referred to as an Auxiliary Classifier Generative Adversarial 
Network (ACGAN). The application to this method is that the different redshift data available from N-body simulations can 
be classified and given to an ACGAN. The generator component of this model can then create samples at different redshifts. 
These samples could be tessellated together to form a larger distribution across a continuous range of  redshifts. A very 
similar method known as a Conditional-GAN (CGAN)~\cite{cgan} is known to work well at the same task.

In the ACGAN framework each generated sample has a corresponding label $c \sim p_c$ and noise vector $\mathbf{z} \sim p_z$ 
which make the generator create a sample $G(c, \mathbf{z})$. The discriminator $D$ gives probability distributions $P(S|X)$ 
and $P(C|X) = D(X)$ over the sources of the samples $S$ and class labels $C$. The objective function is made of two 
independent parts; the likelihood of a correct source $L_S$ and the log-likelihood of the correct class $L_C$

\begin{eqnarray}
    L_S &=& \EE [\log P(S = \textup{real} | X_{real})] \nonumber \\
        &+& \EE [\log P(S = \textup{fake} | X_{fake})] \\ 
        & & \nonumber \\ 
    L_C &=& \EE [\log P(C = c | X_{real})] \nonumber \\
        &+& \EE [\log P(C = c | X_{fake})]
\end{eqnarray}

In this case $D$ tries to maximize $L_S + L_C$ and $G$ tries to minimise $L_C - L_S$.

@@

\begin{figure}
\includegraphics[width=\columnwidth]{figures/diagrams/ACGAN_diagram.png}
\centering
\caption{Schematic of the ACGAN framework; including discriminator and generator networks, the latent-space 
         vector~$\mathbf{z}$, the class label~$\mathbf{c}$, image~$\mathbf{x}$ and the discriminator 
         classifications of the class and source of an image $p_{\mathbf{x}}$ and $p_\mathbf{c}$.}
\label{fig:GAN_diagram}
\end{figure}

@@

% \subsection{Fourier Transform ACGAN (FT-ACGAN)}
% write this if I get it to work

@@

The galaxy catalogs containing physical properties of galaxies in the Millennium simulation were derived with the 
Semi-Analytic Galaxy Evolution (SAGE) codebase by Croton et al. 2016~\cite{Croton2016} that improves on the earlier 
model of Croton et al. 2006~\cite{Croton2006}. The SAGE catalogs were sourced from the Theoretical Astrophysical 
Observatory (TAO)~\cite{TAO}. SAGE modelling adds baryonic processes into N-body simulations of halos after the 
simulations have been run. The baryonic particle properties are from information in the simulation run such as mass, 
size, spin and their merger history. SAGE models account for different mechanisms that affect the evolutions of 
galaxies and therefore the evolution of cosmological structure. The N-body particle positions of the Millennium-SAGE 
data were taken at snapshots at discrete redshifts $z=\{0.00, 1.07, 2.07, 3.06, 4.18, 5.28\}$ to give the training 
data for the ACGAN.

@@

\subsection{Extraction of training data} 
% TRAINING DATA AND PREPROCESSING
Position vector components for the centres of cube samples in the simulation volume were drawn from a uniform distribution 
along each axis of each redshift-snapshot. If the addition or subtraction of half the side length of the cube to any of 
the components resulted in a point outside of the simulation volume, the point was discarded. The galaxy cube samples were 
obtained by sorting the position vectors of galaxies in the simulation that were within the cube volume. Each point inside 
the cube was recorded in an array. The data was augmented by rotating the simulation box through random Euler angles to 
extend the number of possible unique cubes. From each simulation volume of $512 \: h^{-1} \,\textup{Mpc}$ ensembles of 64 
array samples of size $200 \: h^{-1} \,\textup{Mpc}$ were selected. The galaxy positions in each array for every cube were 
scaled to the range $[-1,+1]$. 3D-histograms with 32 bins per axis were made on each cube across the same range for each 
axis. The densities in each bin cell were also scaled to this range to fit the output ranges of the GAN models. 
% After testing different sizes, a cube size of $160 \: h^{-1} \,\textup{Mpc}$ was found to balance the selection of unique 
% samples in the simulation volume against featuring the same features in the simulation structure.

\begin{figure}
\includegraphics[width=\columnwidth]{figures/diagrams/scale_fig.png}% scaling2.png
\centering
\caption{The application of the scaling $s$ to a 3D-histogram training sample. The cube on the left shows an 
image $\mathbf{x}$ and on the right is a scaled image $s(\mathbf{x})$.}
\label{fig:scaling}
\end{figure}

@@

\subsection{Initial data transformations} 
The 3D-histograms have pixel densities that span across many orders of magnitude from voids to dense clusters of galaxies 
for each redshift value. The origin of GANs from image generation tasks has optimised the various architectures present in 
the literature for the general statistical structure of these images - the pixel densities in natural images are typically 
more uniformly distributed over the same scales. To help the GAN in generating sparsely distributed densities a transform, 
first utilised in~\cite{web_gan}, was used on the histograms to enhance the contrast between the density features. The 
transform from the original histogram $\mathbf{x}$ to the scaled histogram $s(\mathbf{x})$ is

\begin{equation}
    s(\mathbf{x}) = \frac{\mathbf{x} - a}{\mathbf{x} + a}
\end{equation}

where $a$ is a free parameter that depends on the sparsity of the histogram densities. This is applied element-wise across 
the training data for all redshifts $z$. The value of $a=1.25$ was found to be optimal for the 
$200 \:  \: h^{-1} \,\textup{Mpc}$ histograms at all redshifts $z$. This parameter remains fixed during training. The value 
of $a$ controls the median pixel density of the scaled image. Figure \ref{fig:scaling} shows the difference between scaled 
images $s(\mathbf{x})$ and the original images $\mathbf{x}$. The inverse transformation $s^{-1}(\mathbf{x})$ can be applied 
to obtain the original image. 
% A logarithmic scaling was used originally, but this did not provide good results with the GAN architecture that used 
% the hyperbolic tangent ($\tanh$) activation because the logarithm of a negative number is undefined. The models using 
% $\tanh$ activations achieved much better results than models without these activations.

@@

\textit{Architecture} --- 
Table~\ref{table:acgan_archi} of Appendix~\ref{appendix:acgan_setup} shows the architectures of the ACGAN discriminator 
and generator models. The hyper-parameters used in the ACGAN are also given in Table~\ref{table:acgan_params} of 
Appendix~\ref{appendix:acgan_setup}. Together the discriminator and generator models had around $2.5\times10^8$ trainable 
parameters across one linear layer and three convolutional layers for the generator model and three convolutional layers 
for the discriminator model. \\

@@

The ACGAN implementation was limited by the GPU memory capacity - the parameters and architectures together with the 
sample batch sizes meant that the GPU was easily saturated. Deeper models with more convolutional units were implemented 
but these tended to stall the GPU. A training loop for $10^5$ epochs for the $200 \: h^{-1} \: \textup{Mpc}$ cubes 
formatted to $32 \times 32 \times 32$ pixel histograms took approximately 144 hours on an NVIDIA Tesla V100 GPU with 
32GB capacity.

@@

% {\setlength{\parindent}{0cm}
% \textit{Redshift interpolation} --- 
% To observe the A fixed point in the latent space is chosen and 
% }
\subsubsection{Redshift interpolation}\label{methods:z_interp}
% talk about what this is in our method, what comes out of doing it 
The latent space of a generator in a typical GAN framework stores a compressed representation of the statistical structure
of the distribution of the real data. By choosing different points in this space and supplying them to the generator, 
samples can be derived with different characteristics that have been extracted from the training process of the 
discriminator and generator models.

An array of latent-space vectors and class labels of a given redshift are passed to the trained ACGAN generator to create
a collection of cube samples. These are placed together face-to-face to create larger samples of cosmic structure. A 
discussion on the viability of this approach is left to Section~\ref{sec:results}.

In the specific case of interpolating through redshift $z$ values, a single point in the latent-space was chosen. This was 
repeatedly supplied to the generator with a class label $c$ that was changed by a small amount $\textup{d}c \sim 0.01$. 
This demonstrates the understanding of the generative model on the relationship between realisations of cosmic structure 
at different values of $z$. The results of this are left to Section~\ref{sec:results}.
% might be bad because of not enough training data, normalisation, parameter capacity in G, z-dim value

@@

\subsubsection{Latent space navigation}\label{methods:z_navig}
% show the maths behind it, the general method of doing it. See https://arxiv.org/pdf/1609.04468.pdf, % (Larsen et al., 2016).
In all non-trivial cases, the number of parameters available to the generative model of a GAN through its architecture is 
less than the total parameters that constitute the training set. This means the discriminator and generator must find 
condensed representations of the training data. The generator model is sampled from a set of latent variables in the 
latent space that constitute the latent-vectors $\mathbf{z}$. Vector space arithmetic with points in the latent space 
allows interpolation between tenable examples of learned representations of the training data the GAN is exposed to in 
training. 
% With the inherent structuring of the latent space by GANs, these operations can allow complex

Interpolation is the transit between two distinct points in the latent space. In this work, spherical interpolation is used
as opposed to linear interpolation. The latter being a over-simplistic approach as the latent space of the model trained in
this work had 256 dimensions with a Gaussian prior $\mathcal{N}(0,1)$. Spherical interpolation~\cite{spherical_interp} uses 
a great circle arc on the N-dimensional hypersphere in the latent space, with a path given by

where a discrete set of points between points $q_1$ and $q_2$ are passed to the generator model to create samples~with.

@@

\subsubsection{Building larger volumes from generated samples}\label{methods:volume_building}
% say how the larger volumes are made from smaller generated cubes, talk about 
A selection of cubes were generated with a batch of $N$ latent vectors $\mathbf{Z}=\{\mathbf{z}_i, ..., \mathbf{z}_N\}$ 
drawn from the same prior $\mathcal{N}(0,1)$ as the generative model in training. These are placed at points on a cubic 
mesh to create a larger box of generated cubes. This was repeated for side-lengths 7, 9 and 11 (in multiples of the initial 
generated cube side-lengths).

@@

\subsection{Comparison of distributions with Kullback-Leibler Divergences}
Generative Adversarial Networks reduce the difference of the model probability distribution between the generated samples 
$p_{model}$ and the probability distribution of the training data samples $p_{data}$. This reduction quantifies how well 
$p_{model}$ estimates $p_{data}$. An estimate on the suitability of this replacement was made using a comparison with 
different ensembles of samples. The ensembles were collected from the training, generated and uniform distributions of 
histograms denoted $t$, $g$ and $u$ at each redshift $z$. The uniform distribution $u$ was a set of histograms with pixel 
densities drawn from a uniform distribution $\mathcal{U}(-1,1)$ so the samples are not discriminated for a redshift value 
in any comparison. The comparison of sample ensembles tested how well the generator made the discriminator associate the 
real and generated samples to the same distribution. The Kullback-Leibler (KL) divergence and its relevance to the 
comparisons is defined in Appendix~\ref{appendix:kl_div}.

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/diagrams/kernel_fig.png}
\centering
\caption{A histogram down-sampled with three kernels of size $k=\{1, 2, 4\}$ (from left to right).}
\label{fig:kernel_fig}
\end{figure}

@@

\subsubsection{Deriving the KL comparative diagnostic}
There were 32 samples in the ensembles from the distributions $t$, $g$ and $u$ that contained 64 samples each for each $z$ 
value. Each histogram in the ensemble was down-sampled by iterating a kernel of size $k$ over each histogram and summing 
the densities. The summed densities were divided by the number of bins in the kernel after down-sampling. This is equivalent 
to the initial histograms having a lower sampling with larger bins. After an ensemble is down-sampled, the ensembles of two 
distributions are compared. The samples are individually flattened to `histogram-vectors' with $(32/k)^3$ components. 
This is shown in Figure~\ref{fig:kernel_fig}. One-dimensional histograms with two bins were made for each component in the 
histogram-vectors. The two bins represented `high' and `low' pixel densities and the boundary between them was the median 
of the training data pixel densities. This is because using the median meant that it was not assumed that the pixel density 
distributions were symmetric. The histogram-vectors were then collapsed over the ensemble and normalised to sum to one as 
a probability distribution. The vector was then sampling the probability density function of pixel density. This process 
is illustrated in Figure \ref{fig:klstat}.

@@

\subsection{Generated images}
% show generated images at various redshifts, compare with training data
Figures~\ref{fig:t_z_imgs} and ~\ref{fig:g_z_imgs} show separate collections of simulated cosmic web and GAN-generated 
cosmic web for each class of redshift in the Millennium-SAGE training data. Across each value of redshift the ACGAN 
demonstrates its ability to evaluate each point in the latent space to a strong representation

\begin{figure*}[hbt!]
\includegraphics[width=15cm]{figures/cubes/classes3d.jpg}
\centering
\caption{A collection of training data cubes extracted from the Millennium-SAGE catalogs. Each row corresponds to the 
         redshift value given above each cube. The scaling $s(\mathbf{x})$ is applied to each cube $\mathbf{x}$.}
\label{fig:t_z_imgs}
\end{figure*}

\begin{figure*}[hbt!]
\includegraphics[width=15cm]{figures/cubes/g_classes3d.png}
\centering
\caption{A collection of generated cubes extracted created by the generator model. Each row corresponds to the 
         redshift value given above each cube. The scaling $s(\mathbf{x})$ is applied to each cube $\mathbf{x}$.}
\label{fig:g_z_imgs}
\end{figure*}

@@

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/mill_box_3.png}
\centering
\caption{A box made of generated cubes at redshift $z=3.19$.}
\label{fig:kernel_fig}
\end{figure}

\begin{figure}[!ht]%[hbt!]
\includegraphics[width=\columnwidth]{figures/cubes/gen_box_9_z3.png}
\centering
\caption{A box made of generated cubes at redshift $z=3.00$.}
\label{fig:kernel_fig}
\end{figure}

@@

\subsection{Loss and $D_{KL}$ curves}
% separation of D and G losses because of number of images shown to models in one training step

% LOSSES AND ACCURACIES
Figure~\ref{fig:acgan_losscurve} shows the ACGAN objectives and accuracies over a run of training. The behaviours of the 
models over the training are apparent in these curves. The sharp jump at the start over the first few 100 epochs or so 
shows the initial learning for both models that are new to their tasks.

There is a sharp jump in the losses of both models, that is highest for the discriminator, which implies the generator 
attempted a new strategy for its objective.

There are small fluctuations in the accuracies of either model as the generator makes smaller adjustments 
relative to the initial training, until just after ?$\times $? epochs
